Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-02-11 VegOzone.tex
Confirm name of bertula species. Add caption to figure. Remove more obsolete text.

\subsection{Model parameters}
Here, we will derive bespoke parameterizations for the main vegetation types (Norway spruce, birch, perennial grassland) surrounding the NIBIO Environment Centre Svanhovd. We roughly estimated land use types for a $1\times 1\,\unit{km}$ patch centered at Svanhovd from areal pictures at kartverket.no: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.

A comprehensive list of model parameters used as starting point in this study can be found in Supplement~Table~S1. These default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual}. Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT are estimated as follows. Evergreen needle leaf trees are photosynthetic active already at low temperatures \citep{TB:Kolari2007}. This leads necessarily to an increase in observable net photosynthesis. Hence, we use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine onset and end of growing season. Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}). By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{day~of~year}$ for 2018 and day 106 in 2019. $A_\text{end}$ amounts to day 261 and 274, respectively, and shall be used for all vegetation alike. {\bf TODO: Compare with latitudinal model?} The resulting growing season for Norway spruce in 2019 was one month longer then in 2018. For estimating $A_\text{start}$ of birch tree, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}). We find 129 and 130, respectively {\bf Accuracy?}. For perennial grassland, we assume a latency period of about $1\,\unit{month}$ after snow melt. Data for first snow-free day is taken from the weather station at Øvre Neiden in Sør-Varanger. In both years, the first snow-free day incidents with the estimated date of bud burst. As leafs of birches in northern Fennoscandia \emph{Betula nana} {\bf (TODO: Right name?)} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$. \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
%To localize leaf area index (LAI), we roughly characterize land use for the $1\times 1\,\unit{km}$ patch centered at Svanhovd from areal pictures at kartverket.no: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban. We normalize LAI to deciduous trees based on standard $\mathrm{DO_3SE}$ parameterization: $f_\text{con}=5/4\cdot f_\text{dec}$, $f_\text{grass}=7/8\cdot f_\text{dec}$. With these, we re-scale the average maximum LAI from MODIS ($\left<f\right> = 2.7$). This results in 3.75, 3, and 2.625 for Norway spruce, birch, and grassland, respectively.

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
  \includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
\caption{TEXT}
\label{fig:f_temp_grassland}
\end{figure}

@@

Except for perennial grassland, model integrations with climatological forcings show higher $\mathrm{POD_y}$ than 2018. This can be explained by means of non linearity in modeled stomatal conductance (Eq.~(\ref{eq:stomatal})) in Eq.~(\ref{eq:pod}). The climatology in $\mathrm{POD_y}$ therefore represents rather an ideal year for growth then a true climatology. To establish a true climatology, it would be necessary to run the $\mathrm{DO_3SE}$ model for 10--20 years with respective observed phenology as well as meteorological and ozone forcing.

Based on this interpretation of climatology in $\mathrm{POD_y}$, we can learn more about potential ozone damage. The time derivative $\frac{\mathrm{\Delta POD_y}}{\Delta t}$ displays very different features for each PFT and meteorological forcing. Although Norway spruce shows a low interannual variation, $\frac{\mathrm{\Delta POD_y}}{\Delta t}$ (Fig.~\ref{fig:pody_rel}) indicates an extended period of higher than average ozone uptake in spring 2018 (steep slope), while ozone uptake in June/July was anomalously low in 2019 (shallow slope) but high in fall. This is most prominently visible in Fig.~\ref{fig:pody}h) for perennial grassland, where consecutive periods of low $G_\text{sto}^\text{leaf}$ manifest in three plateaus in $\mathrm{POD_y}$ throughout the season (1: $184-198\,\unit{doy}$, 2: $209-228\,\unit{doy}$, 3: $235-240\,\unit{doy}$). An episode with anomalous high ozone in mid June ($194\,\unit{doy}$) can be identified but does not amount to unusually high ozone uptake on an annual basis.

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{DO3SE_results_rel}
  \caption{Daily averaged $\mathrm{\Delta POD_y}/\Delta t$ from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October. Error bars/-bands indicate $1 \sigma$ standard deviation. (a) Birch; (b) Norway spruce; (c) perennial grassland.}
  \label{fig:pody_rel}
\end{figure}

>>>>

\subsection{Model parameters}
We roughly estimated land use types for an approximately $1\times 1\,\unit{km}$ patch centered at the NIBIO Environment Centre Svanhovd from Fig.~\ref{fig:svanhovd_research_station}: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
In the following, we derive bespoke parameterizations for the main vegetation types, which we refer to as Norway spruce, birch, perennial grassland.

A comprehensive list of model parameters used as starting point in this study can be found in Supplement~Table~S1. These default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual}. Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT are estimated as follows. Evergreen needle leaf trees are photosynthetic active already at low temperatures \citep{TB:Kolari2007}. This leads necessarily to an increase in observable net photosynthesis. Hence, we use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine onset and end of growing season. Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}). By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{day~of~year}$ for 2018 and day 106 in 2019. $A_\text{end}$ amounts to day 261 and 274, respectively, and shall be used for all vegetation alike. The resulting growing season for Norway spruce in 2019 was one month longer then in 2018. For estimating $A_\text{start}$ of birch tree, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}). We find 129 and 130, respectively. From the latitude model typically used for localizing $\mathrm{DO_3SE}$, we get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ indicating an about $2\,\unit{months}$ longer growing season. For perennial grassland, we assume a latency period of about $1\,\unit{month}$ after snow melt. Data for first snow-free day is taken from the weather station at Øvre Neiden in Sør-Varanger. In both years, the first snow-free day incidents with the estimated date of bud burst. As leafs of birches in northern Fennoscandia \emph{Betula pubescens} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$. \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
  \includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
\caption{Construction of bespoke response functions for grassland. (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown toghether with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability densitiy function - PDF), respectively. Original mapping manual parameterization is shown in comparison as solid line. Note that $Q_0$ has been truncated to $0.006$. PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
\label{fig:f_temp_grassland}
\end{figure}

@@

%Except for perennial grassland, model integrations with climatological forcings show higher $\mathrm{POD_y}$ than 2018. This can be explained by means of non linearity in modeled stomatal conductance (Eq.~(\ref{eq:stomatal})) in Eq.~(\ref{eq:pod}). The climatology in $\mathrm{POD_y}$ therefore represents rather an ideal year for growth then a true climatology. To establish a true climatology, it would be necessary to run the $\mathrm{DO_3SE}$ model for 10--20 years with respective observed phenology as well as meteorological and ozone forcing.

%Based on this interpretation of climatology in $\mathrm{POD_y}$, we can learn more about potential ozone damage. The time derivative $\frac{\mathrm{\Delta POD_y}}{\Delta t}$ displays very different features for each PFT and meteorological forcing. Although Norway spruce shows a low interannual variation, $\frac{\mathrm{\Delta POD_y}}{\Delta t}$ (Fig.~\ref{fig:pody_rel}) indicates an extended period of higher than average ozone uptake in spring 2018 (steep slope), while ozone uptake in June/July was anomalously low in 2019 (shallow slope) but high in fall. This is most prominently visible in Fig.~\ref{fig:pody}h) for perennial grassland, where consecutive periods of low $G_\text{sto}^\text{leaf}$ manifest in three plateaus in $\mathrm{POD_y}$ throughout the season (1: $184-198\,\unit{doy}$, 2: $209-228\,\unit{doy}$, 3: $235-240\,\unit{doy}$). An episode with anomalous high ozone in mid June ($194\,\unit{doy}$) can be identified but does not amount to unusually high ozone uptake on an annual basis.

@@

%\begin{figure}[t]
%  \includegraphics[width=8.3cm]{DO3SE_results_rel}
%  \caption{Daily averaged $\mathrm{\Delta POD_y}/\Delta t$ from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October. Error bars/-bands indicate $1 \sigma$ standard deviation. (a) Birch; (b) Norway spruce; (c) perennial grassland.}
%  \label{fig:pody_rel}
%\end{figure}
